---
title: "Q4 – Airport with Highest Average Departure Delay"
---

# Question

**Find the airport that has the highest average departure delay among all airports.  
Consider 0 minutes delay for flights that departed early.  
Output the airport name, airport code, and the average delay.**

## SQL Query

```sql
SELECT
  OriginAirportName AS airport_name,
  OriginAirportCode AS airport_code,
  AVG(GREATEST(DepDelay, 0)) AS avg_delay
FROM al_perf
GROUP BY OriginAirportName, OriginAirportCode
ORDER BY avg_delay DESC
LIMIT 1;
```
## Output 

The output of this SQL query is saved as `Question 4 output.csv`, and is provided below.

```{r message=FALSE, warning=FALSE}
library(tidyverse)

q4 <- read_csv("data/Question 4 output.csv")
q4
```

## Visualization

```{r}
q4 %>%
ggplot(aes(x = airport_code, y = avg_delay)) +
geom_col(fill = "#4C72B0") +
labs(
title = "Airport with the Highest Average Departure Delay",
x = "Airport Code",
y = "Average Delay (minutes)"
) +
theme_minimal(base_size = 14)
```

The visual for this output is not going to provide a lot of information as the question is only asking for the airport with the highest average departure delay. As we can see, the answer is Bingham airport in New York.